 
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Fall 2020 - Math 421 - Midterm"
---

-------

# Instruction

The midterm has two components: the Rmarkdown notebook (html) and the presentation.  The notebook and the presentation file are due Nov 26. We will do the presentation in class 

**The notebook:** The notebook should be created using `rmarkdown` (like other assignments). The notebook should have a title. 


**The Presentation:** Present your results in 5-10 minutes. To make the presentation using Rmarkdown, do the follows: 

    - In Rstudio -> File -> New File -> R markdown
    
    - In the left panel, click to Presentation -> Click OK
    
    - Now you have an Rmarkdown that can be knitted to be a html presentation 

**What to present**:

  - Present Part 2 - Visualization
  
  - Present Question Question 4, 5 and 6 in Part 3.  
  
  - Present any errors/challenges you run into and how you fix/overcome them. 

**Data:**  

The data for the mid-term project is the Rhode Island Department of Health Hospital Discharge Data.  Each row of the data presents a patient. 

Link: https://drive.google.com/open?id=15QNBf6YYKocK2nNIfpKDer58kQnCPNZJ 

-------

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data. 

```{r}
library(haven)
df <- read_sas("C:/Users/student/Documents/GitHub/emmaschwartz1.github.io/hdd0318cy.sas7bdat")
```

3. Filter the data to have only patients of the year 2018 (`yod==2018`)
```{r}
library(tidyverse)
df <- df %>% filter(yod==18)
```

4. Select to work with only following variables: 

```{r, eval=FALSE}
                      "yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day"
```
```{r}
df <- df %>% select(yod, payfix,pay_ub92,age,  
                      sex,raceethn,provider,moa, 
                      yoa,mod,admtype, asource , 
                      preopday ,los, service , icu,ccu,    
                      dispub92,payer ,drg,trandb, 
                      randbg,randbs,orr, anes,seq,   
                      lab,dtest, ther,blood,phar, 
                      other,patcon,bwght,total,tot ,  
                      ecodub92,b_wt,pt_state,diag_adm,ancilar ,
                      campus,er_fee,er_chrg,er_mode,obs_chrg,
                      obs_hour,psycchrg,nicu_day)
```
 

*Notice*:  You may want to save the current data to your computer for easy access later.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

```{r}
write_csv(df, "midterm.csv")
```

5. What are variables that have missing values?

```{r}
colSums(is.na(df))
```

6. Remove all variables with missing values
```{r}
df$payfix <- NULL
df$preopday <- NULL
df$obs_hour <- NULL
df$nicu_day <- NULL
```

7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?

```{r}
#month of admission is MOA and month of discharge is MOD
```

8. Which month admitted the most number of patients? Which month admitted the most number of male patients?

```{r}
table1 <- table(df$moa)
table
#October has the greatest number of admitted patients

df1 <- df %>% filter(sex=="1") 
table2 <- table(df1$moa)
table2
#October has the greatest number of admitted male patients
```

9. Which month has the most number of teenage female patients?

```{r}
df1 <- df %>% filter(sex=="2") %>%
  filter(age > 12 & age < 20)
table3 <-table(df1$moa)
table3
#March has the greatest number of admitted teenage female patients
```

10. Which provider has the most number of female patients in October? 

```{r}
df1 <- df %>% filter(moa==10 & sex =="2")
table4 <- table(df1$provider)
table4
#Provider 7205 (Rhode Island Hospital) has the greatest number of female patients in October
```

11. Is female patients older than male patients, on average? 

```{r}
df %>% group_by(sex) %>%
  summarize(mean(age))
#No, male patients are older than female patients on average by a little more than half a year
```

12. Calculate the average age of patients by months. Which month has the oldest patients on average age?

```{r}
df %>% group_by(moa) %>%
  summarize(mean(age))
#January has the oldest patients on average
```

13. What is the name of the provider that has the highest total charge?

```{r}
df %>% group_by(provider) %>%
  summarise(max(tot))
#Provider 7214 (Women and Infants) has the highest total charge
```

14. What is the name of the provider that has the least total charge for teenage male on average?

```{r}
df %>% filter(sex=="1" & age <20 & age >12) %>%
  group_by(provider) %>%
  summarise(mean(tot))
#Provider 7206 (Roger Williams) has the least total charge for teenage male on average
```

15. Calculate the length of stays by races.  Which race has the longest length of stays on average?

```{r}
df %>% group_by(raceethn) %>%
  summarise(mean(los))
#Unknown has the longest length of stays on average, but American Indian, not Hispanic has the longest other than that
```

16. On average, how much a 20 year-old male white get charged for staying 1 day?

```{r}
df %>% filter(sex=="1" & age==20 & raceethn==1 & los==1) %>%
  summarise(mean(tot))
#15,319.56 on average
```

-------

## II. Data Visualization

Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis

```{r}
library(ggplot2)

df %>% filter(age < 13) %>%
  ggplot() +
  geom_bar(mapping = aes(x = moa), fill = "#FF91CA")+
  labs(x = 'Month of Admission',
       y = 'Number of Patients',
       title = 'Number of Children Admitted By Month',
       caption = 'Less children under 13 were admitted in the summer months, but overall there is not a strong pattern') +
  scale_x_continuous(breaks = seq(1,12, by=1)) +
  ggsave("plot1.png")
  
```
```{r}
df %>% 
  ggplot() +
  geom_point(mapping = aes(x = los, y = tot))+
  geom_smooth(mapping = aes(x = los, y = tot), color = "#FF91CA")+
  labs(x = 'Length of Stay in Days',
       y = 'Total Charge in Dollars',
       title = 'Length of Stay versus Total Charge',
       caption = 'They are positively correlated, but very long or very expensive stays are uncommon')+
  ggsave("plot2.png")
```

```{r}
df %>%
  ggplot()+
  geom_bar(mapping = aes(x = provider), fill = "#FF91CA")+
  labs(x = 'Provider',
       y = 'Number of Patients',
       title = 'Number of Patients per Provider',
       caption = 'Rhode Island Hospital has the most patients by far; Bradley has the least')+
  ggsave("plot3.png")
```

```{r}
df %>% filter(provider==7205) %>%
  group_by(age) %>%
  summarise(average_los = mean(los)) %>%
  ggplot() +
  geom_point(mapping = aes(x = age, y = average_los))+
  geom_smooth(mapping = aes(x = age, y = average_los), color = "#FF91CA")+
  labs(x = 'Age',
       y = 'Average Length of Stay',
       title = 'Length of Stay by Age at Rhode Island Hospital',
       caption = 'There is less variability of length of stay in patients over 50 years old')+
  ggsave("plot4.png")
  
```

```{r}
df %>% filter(provider==7206 | provider==7209 | provider == 7216) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = age), bins = 10, color = "#FF91CA", fill = "#FFA5C2") +
  facet_wrap(~provider)+
  labs(x = "Age",
       y = "Number of Patients",
       title = "Patient Ages at Roger Williams, South County, and Butler",
       caption = "South County appears to have a lot of infant patients; Roger Williams is the most symmetrical around the mean")+
  ggsave("plot5.png")
```

```{r}
df %>% ggplot() +
  geom_bar(mapping = aes(x = moa, fill = sex), position = 'dodge')+
  labs(x = "Month",
       y = "Number of Patients",
       title = "Number of Males vs Females Admitted Every Month",
       caption = "There are consistently more females admitted each month")+
  ggsave("plot6.png")
```

```{r}
df %>% filter(age>65) %>%
  group_by(age) %>%
  summarise(average_los = mean(los)) %>%
  ggplot() +
  geom_point(mapping = aes(x = age, y = average_los))+
  geom_smooth(mapping = aes(x = age, y = average_los), color = "#FF91CA")+
  labs(x = "Age",
       y = "Averarge Length of Stay",
       title = "Average Length of Stay for Senior Citizens",
       caption = "Looking at senior citizens, as they get older, their length of stay shortens")+
  ggsave("plot7.png")
```

```{r}
df %>% 
  ggplot() +
  geom_bar(mapping = aes(x = provider, fill = sex), position = 'dodge')+
  labs(x = "Provider",
       y = "Number of Patients",
       title = "Amount of Male vs Female Patients by Provider",
       caption = "Provider 7214 is Women and Infants, which explains the large gap in gender there")+
  ggsave("plot8.png")
```

```{r}
df %>%
  group_by(raceethn) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot() +
  geom_col(mapping = aes(x = raceethn, y = average_tot), fill = "#FF91CA")+
  labs(x = "Race/Ethnicity",
       y = "Average Total Charge",
       title = "Average Total Charge by Race",
       caption = "Asian has the lowest average total charge, besides unknown")+
  ggsave("plot9.png")
  
```

```{r}
df %>% group_by(age) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot() +
  geom_line(mapping = aes(x = age, y = average_tot), color = "#FF91CA")+
  labs(x = "Age",
       y = "Average Total Charge",
       title = "Average Total Charge by Age",
       caption = "Total cost appears to be highest for teenagers")+
  ggsave("plot10.png")
```

2. Make an animation

```{r}
library(gganimate)
df %>% group_by(los) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot(aes(x = los, y = average_tot)) +
  geom_line()+
  geom_point(size = 3)+
  transition_reveal(los)+
  labs(x = "Length of Stay",
       y = "Average Total Charge",
       title = "Average Total Charge by Length of Stay",
       caption = "While positively correlated, there are a lot of spikes throughout")
```

-------

## III. Predictive Models

Continue with the data from part I. Use the follows as the target and input variables: 

*Target Variable*: Create the target variable taking value of 

  - `low` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high` otherwise. 

*Input Variables*:

  - "age","sex","raceethn","provider","moa","mod","admtype","campus", 'los'
  
-------

1. Use `filter` function to filter out rows where `raceethn==''` or `admtype==''`. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 

```{r}
library(tidyverse)

df$target <- case_when(
  df$tot < median(df$tot) ~ 'low',
  TRUE ~ 'high'
)

df <- df %>% select(age, sex, raceethn, provider, moa, mod, admtype, campus, los, target)
df <- df %>% drop_na(raceethn, admtype)


library(caret)

set.seed(2020)
splitIndex <- createDataPartition(df$target, p = 0.90,
                                  list = FALSE)
df_train <- df[splitIndex, ]
df_test <- df[-splitIndex, ]


```

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 

```{r}
library(rpart)
tree_model <- rpart(target ~., data = df_train,
                    control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(tree_model)

tree_model$variable.importance
barplot(tree_model$variable.importance)

```

3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 3 different models (i.e. three different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model. 

```{r}
tuneGrid = expand.grid(maxdepth = 2:9)
trControl = trainControl(method = "cv",
                         number = 5)

tree <- train(target~., data = df_train,
                        method = "rpart2",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(tree)


tuneGrid = expand.grid(nIter = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)

library(caret)

logit <- train(target~., data = df_train,
                        method = "LogitBoost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(logit)

  
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(maxdepth = 2:5,
                       mstop = 3)

blackboost <- train(target~., data = df_train,
                        method = "blackboost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(blackboost)
```

4. Plot the comparison of the models in 3. 

```{r}
results <- resamples(list(tree = tree,
                          logit = logit,
                          blackboost = blackboost))
bwplot(results)
```

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 

```{r}
pred <- predict(logit, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "low")
cm$overall[1]

```

6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 

```{r}
library(tidyverse)

df <- read_csv("midterm.csv")
df$target <- case_when(
  df$los < 7 ~ 'short',
  TRUE ~ 'long'
)

df <- df %>% select(age, sex, raceethn, provider, moa, mod, admtype, campus, target)
df <- df %>% drop_na(raceethn, admtype)


library(caret)

set.seed(2020)
splitIndex <- createDataPartition(df$target, p = 0.90,
                                  list = FALSE)
df_train <- df[splitIndex, ]
df_test <- df[-splitIndex, ]

```

```{r}
library(rpart)
tree_model <- rpart(target ~., data = df_train,
                    control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(tree_model)

tree_model$variable.importance
barplot(tree_model$variable.importance)
```

```{r}
tuneGrid = expand.grid(maxdepth = 2:8)
trControl = trainControl(method = "cv",
                         number = 5)

tree <- train(target~., data = df_train,
                        method = "rpart2",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(tree)


tuneGrid = expand.grid(nIter = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)

library(caret)

logit <- train(target~., data = df_train,
                        method = "LogitBoost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(logit)

  
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(maxdepth = 2:5,
                       mstop = 3)

blackboost <- train(target~., data = df_train,
                        method = "blackboost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(blackboost)
```

```{r}
results <- resamples(list(tree = tree,
                          logit = logit,
                          blackboost = blackboost))
bwplot(results)
```

```{r}
pred <- predict(logit, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "short")
cm$overall[1]
```

-------