
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(tidyverse)
library(ggplot2)
library(gganimate)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df %>% filter(Date_reported == "2020-10-08") %>%
  arrange(-Cumulative_deaths)
```


2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths

```{r}
library(lubridate)
df$date <- ymd(df$Date_reported)
df %>% filter(Country=='United States of America'|Country=='Brazil'|Country=='India') %>%
  ggplot() +
  geom_line(mapping = aes(y = Cumulative_deaths, x = date, color = Country)) +
  transition_reveal(date) +
  labs(title = "Cumulative Deaths Top Countries", x = "Date", y = "Number of deaths")
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

```{r}
df$death_per_cases <- df$Cumulative_deaths / df$Cumulative_cases
df %>% arrange(-death_per_cases)

#San Marino, Gambia, and Nicaragua have the highest deaths per cases.
```

4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 

```{r}
library(lubridate)
df$date <- ymd(df$Date_reported)
df %>% filter(Country=='United States of America'|Country=='Italy'|Country=='Mexico')  %>%
  ggplot() +
  geom_line(mapping = aes(x = date, y = death_per_cases, color = Country))+
  transition_reveal(date) +
  labs(title = "Deaths per Case", x = "Date", y = "Number of deaths per case")
```

