
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

```{r}
library(tidyverse)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```

-------

1. Make a plot use `face_wrap` by one categorical variable (`face_wrap(~variable1)`)
```{r}
library(lubridate)
library(ggplot2)

df$date <-ymd(df$date)
df$month <- month(df$date)

df %>% filter(state=='CT'|state=='NY'|state=='NJ') %>%
  ggplot() +
  geom_col(mapping = aes(x = month, y = positiveIncrease)) +
  facet_wrap(~state) +
  labs(x = 'Month',
       y = 'New Positive Cases',
       title = 'New Cases Among Tri-State Area',  
       caption = 'The three states followed a similar pattern of new cases, just with different amounts')
```


2. Make a plot use `face_wrap` by two categorical variables (`face_wrap(variable1~variable2)`)
```{r}
df %>% filter(state=='CT'|state=='NY'|state=='NJ') %>%
  filter(month==3|month==4|month==5) %>%
  ggplot() +
  geom_point(mapping = aes(x = positiveIncrease, y = deathIncrease)) +
  facet_wrap(state~month) +
  labs(x = 'New Positive Cases',
       y = 'New Deaths',
       title = 'Cases vs. Death among Tri-State Area in Spring',  
       caption = 'New York had the biggest spread in every month')
```


3. Make a plot use `position ='jitter'`. 
```{r}
df %>% filter(state=='NY'|state=='CT'|state=='NJ') %>%
  ggplot() +
  geom_point(mapping = aes(x = date, y = positiveIncrease, color = state), position = 'jitter') +
  labs(x = 'Date',
       y = 'New Positive Cases',
       title = 'New Cases in Tri-State 2020',  
       caption = 'All states in the tri-state area peaked in April, with New York consistently being highest')
```

4. Make a plot use `position ='dodge'`.
```{r}
df %>% filter(state=='NY'|state=='CT'|state=='NJ') %>%
  filter(month==3|month==4|month==5) %>%
  ggplot() +
  geom_col(mapping = aes(x = month, y = positiveIncrease, fill = state), position = 'dodge') +
  labs(x = 'Month',
       y = 'New Positive Cases',
       title = 'New Cases in Tri-State Area this Spring',  
       caption = 'All states in the tri-state area peaked in April')
```

5. Make a plot use `position ='fill'`. 
```{r}
df %>% filter(state=='NY'|state=='CT'|state=='NJ') %>%
  filter(month==3|month==4|month==5) %>%
  ggplot() +
  geom_col(mapping = aes(x = month, y = positiveIncrease, fill = state), position = 'fill') +
  labs(x = 'Month',
       y = 'New Positive Cases',
       title = 'New Cases in Tri-State Area this Spring',  
       caption = 'New York consistently had the highest number of new cases this spring')
```

6. Make a plot use `geom_col`.
```{r}
df %>% filter(state=='CT') %>%
  ggplot(mapping = aes(x = month, y = positiveIncrease)) +
  geom_col() +
  labs(x = 'Month',
       y = 'New Positive Cases',
       title = 'New Cases in Connecticut',  
       caption = 'New cases in Connecticut peaked in April and are on trend to continue decreasing')
```

7. Change the theme of one of your plots. 
```{r}
df %>% filter(state=='CT'|state=='NY'|state=='NJ') %>%
  ggplot() +
  geom_col(mapping = aes(x = month, y = positiveIncrease)) +
  facet_wrap(~state) +
  labs(x = 'Month',
       y = 'New Positive Cases',
       title = 'New Cases Among Tri-State Area',  
       caption = 'The three states followed a similar pattern of new cases, just with different amounts') +
  theme_linedraw() +
  ggsave('plot.png')
```

8. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
[Link to the plot](abc.png)
```





