
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment12.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes

library(caret)
```

- Set seed to be 2020. 
- The target variable is `diabetes`
- Partition the data into 80% training and 20% testing.  

```{r}
set.seed(2020)
splitIndex <- createDataPartition(df$diabetes,p = .80,
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

-------

2. Use cross-validation of 30 folds to tune random forest (method='rf').  What is the `mtry` value that produces the greatest accuracy?
 
```{r}
tuneGrid = expand.grid(mtry = 1:8)
trControl = trainControl(method = "cv",
                         number = 30)

forest_cv <- train(diabetes~., data = df_train,
                   method = "rf",
                   trControl = trControl,
                   tuneGrid = tuneGrid)
print(forest_cv)

#The final value used for the model is the value with the greatest accuracy
```
 
-------

3. Use cross-validation with of 30 folds to tune random forest (method='ranger').  What are the parameters that produce the greatest accuracy?

```{r}
trControl = trainControl(method = "cv",
                         number = 30)
tuneGrid = expand.grid(mtry = 1:8,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:10))

ranger_cv <- train(diabetes~., data = df_train,
                   method = "ranger",
                   trControl = trControl,
                   tuneGrid = tuneGrid)
plot(ranger_cv)
```

-------

4. Go to https://topepo.github.io/caret/available-models.html and pick a classification model.  Tune the classification model using cross-validation of 30 folds. 

```{r}
tuneGrid = expand.grid(mtry = 1:8)
trControl = trainControl(method = "cv",
                         number = 30)

parallel_cv <- train(diabetes~., data = df_train,
                   method = "parRF",
                   trControl = trControl,
                   tuneGrid = tuneGrid)

print(parallel_cv)
```

-------

5. Compare the three models in question 2, 3, and 4 to select the final model.  Evaluate the accuracy of the final model on the test data. 

```{r}
results <- resamples(list(forest = forest_cv,
                          ranger = ranger_cv,
                          parallel = parallel_cv))
bwplot(results)

#final model selected is ranger

pred <- predict(ranger_cv, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]


```

