---
title: "Midterm"
author: "Emma Schwartz"
date: "10/25/2020"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

## Children Admitted by Month

```{r, message=FALSE}

library(haven)
df <- read_sas("C:/Users/student/Documents/GitHub/emmaschwartz1.github.io/hdd0318cy.sas7bdat")
library(tidyverse)
df <- df %>% filter(yod==18)
df <- df %>% select(yod, payfix,pay_ub92,age,  
                      sex,raceethn,provider,moa, 
                      yoa,mod,admtype, asource , 
                      preopday ,los, service , icu,ccu,    
                      dispub92,payer ,drg,trandb, 
                      randbg,randbs,orr, anes,seq,   
                      lab,dtest, ther,blood,phar, 
                      other,patcon,bwght,total,tot ,  
                      ecodub92,b_wt,pt_state,diag_adm,ancilar ,
                      campus,er_fee,er_chrg,er_mode,obs_chrg,
                      obs_hour,psycchrg,nicu_day)
df$payfix <- NULL
df$preopday <- NULL
df$obs_hour <- NULL
df$nicu_day <- NULL

library(ggplot2)

df %>% filter(age < 13) %>%
  ggplot() +
  geom_bar(mapping = aes(x = moa), fill = "#FF91CA")+
  labs(x = 'Month of Admission',
       y = 'Number of Patients',
       title = 'Number of Children Admitted By Month',
       caption = 'Less children under 13 were admitted in the summer months, but overall there is not a strong pattern') +
  scale_x_continuous(breaks = seq(1,12, by=1))
```

## Length of Stay and Total Charge

```{r, message=FALSE}
df %>% 
  ggplot() +
  geom_point(mapping = aes(x = los, y = tot))+
  geom_smooth(mapping = aes(x = los, y = tot), color = "#FF91CA")+
  labs(x = 'Length of Stay in Days',
       y = 'Total Charge in Dollars',
       title = 'Length of Stay versus Total Charge',
       caption = 'They are positively correlated, but very long or very expensive stays are uncommon')
```

## Patients by Provider

```{r, message=FALSE}
df %>%
  ggplot()+
  geom_bar(mapping = aes(x = provider), fill = "#FF91CA")+
  labs(x = 'Provider',
       y = 'Number of Patients',
       title = 'Number of Patients per Provider',
       caption = 'Rhode Island Hospital has the most patients by far; Bradley has the least')
```

## Length of Stay by Age

```{r, message=FALSE}
df %>% filter(provider==7205) %>%
  group_by(age) %>%
  summarise(average_los = mean(los)) %>%
  ggplot() +
  geom_point(mapping = aes(x = age, y = average_los))+
  geom_smooth(mapping = aes(x = age, y = average_los), color = "#FF91CA")+
  labs(x = 'Age',
       y = 'Average Length of Stay',
       title = 'Length of Stay by Age at Rhode Island Hospital',
       caption = 'There is less variability of length of stay in patients over 50 years old')
```

## Patient Ages

```{r, message=FALSE}
df %>% filter(provider==7206 | provider==7209 | provider == 7216) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = age), bins = 10, color = "#FF91CA", fill = "#FFA5C2") +
  facet_wrap(~provider)+
  labs(x = "Age",
       y = "Number of Patients",
       title = "Patient Ages at Roger Williams, South County, and Butler",
       caption = "South County appears to have a lot of infant patients; Roger Williams is the most symmetrical around the mean")
```

## Admitted Patients by Gender and Month

```{r, message=FALSE}
df %>% ggplot() +
  geom_bar(mapping = aes(x = moa, fill = sex), position = 'dodge')+
  labs(x = "Month",
       y = "Number of Patients",
       title = "Number of Males vs Females Admitted Every Month",
       caption = "There are consistently more females admitted each month")
```

## Length of Stay by Age

```{r, message=FALSE}
df %>% filter(age>65) %>%
  group_by(age) %>%
  summarise(average_los = mean(los)) %>%
  ggplot() +
  geom_point(mapping = aes(x = age, y = average_los))+
  geom_smooth(mapping = aes(x = age, y = average_los), color = "#FF91CA")+
  labs(x = "Age",
       y = "Averarge Length of Stay",
       title = "Average Length of Stay for Senior Citizens",
       caption = "Looking at senior citizens, as they get older, their length of stay shortens")
```

## Patients by Gender and Provider

```{r, message=FALSE}
df %>% 
  ggplot() +
  geom_bar(mapping = aes(x = provider, fill = sex), position = 'dodge')+
  labs(x = "Provider",
       y = "Number of Patients",
       title = "Amount of Male vs Female Patients by Provider",
       caption = "Provider 7214 is Women and Infants, which explains the large gap in gender there")
```

## Total Charges by Race

```{r, message=FALSE}
df %>%
  group_by(raceethn) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot() +
  geom_col(mapping = aes(x = raceethn, y = average_tot), fill = "#FF91CA")+
  labs(x = "Race/Ethnicity",
       y = "Average Total Charge",
       title = "Average Total Charge by Race",
       caption = "Asian has the lowest average total charge, besides unknown")
```

## Total Charges by Age

```{r, message=FALSE}
df %>% group_by(age) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot() +
  geom_line(mapping = aes(x = age, y = average_tot), color = "#FF91CA")+
  labs(x = "Age",
       y = "Average Total Charge",
       title = "Average Total Charge by Age",
       caption = "Total cost appears to be highest for teenagers")
```

## Animation: Total Charge vs Length of Stay

```{r, message=FALSE}
library(gganimate)
df %>% group_by(los) %>%
  summarise(average_tot = mean(tot)) %>%
  ggplot(aes(x = los, y = average_tot)) +
  geom_line()+
  geom_point(size = 3)+
  transition_reveal(los)+
  labs(x = "Length of Stay",
       y = "Average Total Charge",
       title = "Average Total Charge by Length of Stay",
       caption = "While positively correlated, there are a lot of spikes throughout")
```

## Predictive Models: Model 1

```{r, message=FALSE}

library(tidyverse)

df$target <- case_when(
  df$tot < median(df$tot) ~ 'low',
  TRUE ~ 'high'
)

df <- df %>% select(age, sex, raceethn, provider, moa, mod, admtype, campus, los, target)
df <- df %>% drop_na(raceethn, admtype)


library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df$target, p = 0.90,
                                  list = FALSE)
df_train <- df[splitIndex, ]
df_test <- df[-splitIndex, ]
library(rpart)
tree_model <- rpart(target ~., data = df_train,
                    control = rpart.control(maxdepth = 3))
library(rattle)

tuneGrid = expand.grid(maxdepth = 2:9)
trControl = trainControl(method = "cv",
                         number = 5)

tree <- train(target~., data = df_train,
                        method = "rpart2",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(tree)
```

## Predictive Models: Model 2

```{r, message=FALSE}
tuneGrid = expand.grid(nIter = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)

library(caret)

logit <- train(target~., data = df_train,
                        method = "LogitBoost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(logit)
```

## Predictive Models: Model 3

```{r, message=FALSE}
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(maxdepth = 2:5,
                       mstop = 3)

blackboost <- train(target~., data = df_train,
                        method = "blackboost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(blackboost)
```

## Predictive Models: Comparison

```{r, message=FALSE}
results <- resamples(list(tree = tree,
                          logit = logit,
                          blackboost = blackboost))
bwplot(results)
```

## Predictive Models: Final Selection

```{r, message=FALSE}
pred <- predict(logit, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "low")
cm$overall[1]
```

## More Predictive Models: Model 1

```{r, message=FALSE}
library(tidyverse)

library(haven)
df <- read_sas("C:/Users/student/Documents/GitHub/emmaschwartz1.github.io/hdd0318cy.sas7bdat")
library(tidyverse)
df <- df %>% filter(yod==18)
df <- df %>% select(yod, payfix,pay_ub92,age,  
                      sex,raceethn,provider,moa, 
                      yoa,mod,admtype, asource , 
                      preopday ,los, service , icu,ccu,    
                      dispub92,payer ,drg,trandb, 
                      randbg,randbs,orr, anes,seq,   
                      lab,dtest, ther,blood,phar, 
                      other,patcon,bwght,total,tot ,  
                      ecodub92,b_wt,pt_state,diag_adm,ancilar ,
                      campus,er_fee,er_chrg,er_mode,obs_chrg,
                      obs_hour,psycchrg,nicu_day)
df$payfix <- NULL
df$preopday <- NULL
df$obs_hour <- NULL
df$nicu_day <- NULL
df$target <- case_when(
  df$los < 7 ~ 'short',
  TRUE ~ 'long'
)

df <- df %>% select(age, sex, raceethn, provider, moa, mod, admtype, campus, target)
df <- df %>% drop_na(raceethn, admtype)


library(caret)

set.seed(2020)
splitIndex <- createDataPartition(df$target, p = 0.90,
                                  list = FALSE)
df_train <- df[splitIndex, ]
df_test <- df[-splitIndex, ]
library(rpart)
tree_model <- rpart(target ~., data = df_train,
                    control = rpart.control(maxdepth = 3))
tuneGrid = expand.grid(maxdepth = 2:8)
trControl = trainControl(method = "cv",
                         number = 5)

tree <- train(target~., data = df_train,
                        method = "rpart2",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(tree)
```

## More Predictive Models: Model 2

```{r, message=FALSE}
tuneGrid = expand.grid(nIter = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)

library(caret)

logit <- train(target~., data = df_train,
                        method = "LogitBoost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(logit)
```

## More Predictive Models: Model 3

```{r, message=FALSE}
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(maxdepth = 2:5,
                       mstop = 3)

blackboost <- train(target~., data = df_train,
                        method = "blackboost",
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(blackboost)
```

## More Predictive Models: Comparison

```{r, message=FALSE}
results <- resamples(list(tree = tree,
                          logit = logit,
                          blackboost = blackboost))
bwplot(results)
```

## More Predictive Models: Final Selection

```{r, message=FALSE}
pred <- predict(logit, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "short")
cm$overall[1]
```

## Challenges

- Figuring out what errors mean

- Waiting for code to run

- Deciding how to look at the data
